---
title: "lobster_exploration"
author: "Joe DeCesaro"
date: "8/16/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(janitor)

```

Within your .Rmd, write organized and well-annotated code to do the following:
- Read in and take a look at the data in the data/Lobster_Abundance_All_Years_20210412.csv file. Take note of values that can be considered NA (see metadata) and update your import line to convert those to NA values
- Convert column names to lower snake case
- Convert the data from frequency to case format using dplyr::uncount() on the existing count column. What did this do? Add annotation in your code explaining dplyr::uncount()

```{r}
lobsters <- read_csv(here("data", "Lobster_Abundance_All_Years_20210412.csv"), na = c("-99999", "")) %>% 
  clean_names() %>% 
  uncount(count)

# uncount() is separating the "count" column included in the original data to ensure that no values are excluded
```

Find counts and mean sizes by site & year
- Create a summary table that finds the total counts (see: n()), and mean carapace lengths of lobsters observed in the dataset by site and year.
- Create a ggplot graph of the number of total lobsters observed (y-axis) by year (x-axis) in the study, grouped (either aesthetically or by faceting) by site
```{r}
lobsters_summary <- lobsters %>% 
  group_by(site, year) %>% 
  summarize(length_means = mean(size_mm, na.rm = TRUE), 
            total_counts = n())

ggplot(lobsters_summary, aes(y = total_counts, x = year)) +
  geom_point(aes(color = site))
  
```
Find the proportion of legal lobsters at each site for 2020
- The legal lobster size (carapace length) in California is 79.76 mm.
- Create a subset that only contains lobster data from 2020 (note: this should be from the original data you read in, not the summary table you created above)
- Write code (you can decide how to do this - there are a number of ways) to find the counts of lobsters observed at each site (only using site as the grouping factor) that are above and below the legal limit. Hint: You may want to add a new column legal that contains “yes” or “no” based on the size of the observed lobster (see dplyr::case_when() for a really nice way to do this), then use group_by() %>% summarize(n()) or dplyr::count() to get counts by group within variables
- Create a stacked column graph that shows the proportion of legal and non-legal lobsters at each site. **Hint: create a stacked column graph with geom_col(), then add the argument position = "fill" to convert from a graph of absolute counts to proportions.

Which two sites had the largest proportion of legal lobsters in 2020? Explore the metadata to come up with a hypothesis about why that might be.

```{r}
lobster_2020 <- lobsters %>% 
  filter(year == 2020) %>% 
  mutate(legal = case_when(
    size_mm < 79.76 ~ "no",
    size_mm >= 79.76 ~ "yes")) %>% 
  group_by(site, legal) %>% 
  summarize(count = n())

ggplot(lobster_2020, aes(x = site, y = count)) +
  geom_col(position = "fill", aes(fill = legal))
```
The Naples (NAPL) and Isla Vista (IVEE) sites had the highest proportions of legal lobsters because of their status as marine protected areas.

## Task 3: Random lobster wrangling
Starting with the original lobsters data that you read in as lobsters, complete the following (separately - these are not expected to be done in sequence or anything). You can store each of the outputs as ex_a, ex_b, etc. for the purposes of this task.
### filter() practice

Create and store a subset that only contains lobsters from sites “IVEE”, “CARP” and “NAPL”. Check your output data frame to ensure that only those three sites exist.
```{r}
filter_a <- lobsters %>% 
  filter(site %in% c("NAPL", "IVEE", "CARP"))
unique(filter_a$site)
```
Create a subset that only contains lobsters observed in August.
```{r}
filter_b <- lobsters %>% 
  filter(month == 8)
```

Create a subset with lobsters at Arroyo Quemado (AQUE) OR with a carapace length greater than 70 mm.
```{r}
filter_c <- lobsters %>% 
  filter(site == "AQUE" | size_mm > 70)
```

Create a subset that does NOT include observations from Naples Reef (NAPL)
```{r}
filter_d <- lobsters %>% 
  filter(!site == "NAPL")
```

### group_by() %>% summarize() practice
Find the mean and standard deviation of lobster carapace length, grouped by site.
```{r}
ex_e <- lobsters %>% 
  group_by(site) %>% 
  summarize(mean = mean(size_mm, na.rm = TRUE),
            sd = sd(size_mm, na.rm = TRUE))
```

Find the maximum carapace length by site and month.
```{r}
ex_f <- lobsters %>% 
  group_by(site, month) %>% 
  summarize(max = max(size_mm, na.rm = TRUE))
```

### mutate() practice
Add a new column that contains lobster carapace length converted to centimeters. Check output.
```{r}
ex_g <- lobsters %>% 
  mutate(size_cm = size_mm / 10)
```

Update the site column to all lowercase. Check output.
```{r}
ex_h <- lobsters %>% 
  mutate(site = tolower(site))
```

Convert the area column to a character (not sure why you’d want to do this, but try it anyway). Check output.
```{r}
ex_i <- lobsters %>% 
  mutate(area = as.character(area))
```

### case_when() practice
Use case_when() to add a new column called size_bin that contains “small” if carapace size is <= 70 mm, or “large” if it is greater than 70 mm. Check output.
```{r}
ex_j <- lobsters %>% 
  mutate(size_bin = case_when(
    size_mm <= 70 ~ "small", 
    size_mm > 70 ~ "large"
  ))
```
Use case_when() to add a new column called designation that contains “MPA” if the site is “IVEE” or “NAPL”, and “not MPA” for all other outcomes.
```{r}
ex_k <- lobsters %>% 
  mutate(designation = case_when(
    site == "IVEE" | site == "NAPL" ~ "MPA", 
    TRUE ~ "NOT MPA"))
```

